---
title: "Data Cleaning"
---

> The goal of this project is to clean a dataset for one of my thesis studies. The dataset comes directly as it was downloaded from Qualtrics. I will clean the dataset by removing cases according to the criteria that I preregistered. I'll also combine item scores into variables and combine variables into a composite variable. The overall goal is to create a dataset that is ready for analyses.

```{r}
library(tidyverse)
library(labelled)
library(haven)
qualtrics_data <- read_sav("data/Data from Qualtrics - CSJA Study 2_April 1, 2025_12.34.sav")
```

#n = 176
#Expecting n = 150 valid cases because 150 credits were granted in SONA to respondents
#who proceeded to the end of the survey.

#Per preregistration, using only cases in which participants proceeded to the end of the survey.
```{r}
qualtrics_data <- qualtrics_data %>% 
  filter(Finished == 1)
```
#n = 159

#Removing responses from SONA Admin which were used for technical troubleshooting.
```{r}
qualtrics_data <- qualtrics_data %>% 
  filter(id != "9244")
```
#n = 157

#A technical issue in which the SONA study was not automatically granting participants credit 
#when they completed the Qualtrics survey caused some participants to take the survey more 
#than once. I'll now screen out duplicate survey responses based on the id variable.
```{r}
duplicate_cases <- qualtrics_data %>%
  group_by(id) %>%
  filter(n() > 1) %>%
  arrange(id) %>% 
  select(id, everything())
duplicate_cases
```

#For the first set of duplicates, the response with ResponseId = R_1YeOsLfdtzWYfHD has duration of 12
#seconds and did not proceed further than the eligibility question.
#For the second set of duplicates, both are complete responses. I will remove the response that came later,
#ResponseId = R_7n2BtvDKzUgs0xn.
#Same with the third set. Removing R_1jcEdJAa4cI8FqL.
#Same with the fourth set. Removing R_6dFnR4w70ypPgkp.
#For the fifth set, R_5390VyhkwD8FFRI did not proceed further than the eligibility question.
#For the sixth set, I will remove the later response, R_5fvLGGf4AbBIGX8.
```{r}
qualtrics_data <- qualtrics_data %>%
  filter(!ResponseId %in% c(
    "R_1YeOsLfdtzWYfHD",
    "R_7n2BtvDKzUgs0xn", 
    "R_1jcEdJAa4cI8FqL",
    "R_6dFnR4w70ypPgkp",
    "R_5390VyhkwD8FFRI",
    "R_5fvLGGf4AbBIGX8"
  ))
```

#n = 151.
#Respondent with survey id 26063 answered no the eligibility question (US citizenship)
#and was not permitted to proceed with the survey.
```{r}
qualtrics_data <- qualtrics_data %>%
  filter(id != 26063)
```

#n = 150
#Respondent with survey id 26230 did not answer any questions after the demographic 
#secion of the survey, so they have no usable data and will be removed.
```{r}
qualtrics_data <- qualtrics_data %>%
  filter(id != 26230)
```

#n = 149

#Some items were mistakenly named as reversed items.
```{r}
qualtrics_data <- qualtrics_data %>%
  rename(NFM8 = NFM8r, CSJAS22 = CSJAS22r)
```


#Per preregistration, screening for cases who selected all highest or all lowest
#response options across all surveys for exclusion from analysis.
```{r}
#Calculating each participant's variance for each scale.
qualtrics_data <- qualtrics_data %>%
  mutate(
    respvar_NFM = apply(select(., NFM1r:NFM10), 1, var, na.rm = TRUE),
    respvar_CSJAS = apply(select(., CSJAS1:CSJAS28), 1, var, na.rm = TRUE),
    respvar_RCI = apply(select(., RCI1:RCI10), 1, var, na.rm = TRUE),
    respvar_EXPL = apply(select(., SEP:EL_), 1, var, na.rm = TRUE),
    respvar_ANX = apply(select(., Anx__1:Anx__7), 1, var, na.rm = TRUE),
    respvar_DEP = apply(select(., Dep_1:Dep_13), 1, var, na.rm = TRUE),
    respvar_LS = apply(select(., LS_1:LS_5), 1, var, na.rm = TRUE),
    respvar_PANAS = apply(select(., PANAS__1:PANAS__20), 1, var, na.rm = TRUE)
  )
#Identifying cases with no response variance across all scales.
qualtrics_data %>%
  filter(respvar_NFM == 0,
         respvar_CSJAS == 0,
         respvar_RCI == 0,
         respvar_EXPL == 0,
         respvar_ANX == 0,
         respvar_DEP == 0,
         respvar_LS == 0,
         respvar_PANAS == 0)
```

#Zero cases with no response variance across all quesionnaires. No participants excluded.


#Per preregistration, screening for cases with survey completion times below 3 standard
#deviations of the mean survey completion time for exclusion from analysis.
```{r}
qualtrics_data %>%
  mutate(duration_minutes = Duration__in_seconds_/60) %>% 
  mutate(
    duration_mean = mean(duration_minutes, na.rm = TRUE),
    duration_sd = sd(duration_minutes, na.rm = TRUE),
    z_duration_minutes = (duration_minutes - duration_mean)/duration_sd
  ) %>% 
  filter(z_duration_minutes <= -3)
```

#No participants excluded.

#Storing the variable labels before transformation.
```{r}
var_labels <- var_label(qualtrics_data)
val_labels <- val_labels(qualtrics_data)
```


#Reverse scoring the appropriate NFM items.
```{r}
  #Examining reversed items. (Scored 1-7.)
qualtrics_data %>%
  select(matches("^NFM.*r$")) %>%
  names()
print(qualtrics_data$NFM1r)
print(qualtrics_data$NFM3r)
print(qualtrics_data$NFM7r)
  #Reversing.
qualtrics_data <- qualtrics_data %>%
  mutate(across(matches("^NFM.*r$"), ~8 - .))
  #Verifying reversal.
print(qualtrics_data$NFM1r)
print(qualtrics_data$NFM3r)
print(qualtrics_data$NFM7r)
```

#Creating NFM variable.
```{r}
qualtrics_data <- qualtrics_data %>% 
  mutate(NFM = rowMeans(across(NFM1r:NFM10), na.rm = TRUE))
```

#Reverse scoring the appropriate CSJAS items.
```{r}
  #Examining reversed items. (Scored 1-5.)
qualtrics_data %>%
  select(matches("^CSJAS.*r$")) %>%
  names()
csjas_reversed <- c("CSJAS4r", "CSJAS5r", "CSJAS7r", "CSJAS11r", "CSJAS14r", 
                    "CSJAS15r", "CSJAS18r", "CSJAS20r", "CSJAS21r", "CSJAS23r",
                    "CSJAS25r", "CSJAS26r")
for(item in csjas_reversed) {
  cat("\n--- Item:", item, "---\n")
  print(qualtrics_data[[item]])
}
  #Reversing.
qualtrics_data <- qualtrics_data %>%
  mutate(across(matches("^CSJAS.*r$"), ~6 - .))
  #Confirming reversal.
print(qualtrics_data$CSJAS26r)
```

#Creating the CSJAS variable by combining the 7 final scale items determined in Study 1.
```{r}
qualtrics_data <- qualtrics_data %>% 
  mutate(CSJAS = rowMeans(across(c(CSJAS1, CSJAS2, CSJAS4r, CSJAS10, CSJAS15r, CSJAS17, CSJAS28)), na.rm = TRUE))
```

#Creating the Religious Commitment and Intrapersonal and Interpersonal subscales variables.
```{r}
qualtrics_data <- qualtrics_data %>% 
  mutate(
    Relig = rowMeans(across(RCI1:RCI10), na.rm = TRUE),
    Relig_inter = rowMeans(across(c(RCI2, RCI6, RCI9, RCI10)), na.rm = TRUE),
    Relig_intra = rowMeans(across(c(RCI1, RCI3, RCI4, RCI5, RCI7, RCI8)), na.rm = TRUE)
  )
```

#Creating Anxiety variable.
```{r}
qualtrics_data <- qualtrics_data %>% 
  mutate(
    Anxiety = rowMeans(across(c(Anx__1:Anx__7)), na.rm = TRUE))
```

#Creating Depression variable.
```{r}
qualtrics_data <- qualtrics_data %>% 
  mutate(
    Depression = rowMeans(across(c(Dep_1:Dep_13)), na.rm = TRUE))
```

#Creating Life Satisfaction variable.
```{r}
qualtrics_data <- qualtrics_data %>% 
  mutate(
    LifeSatisf = rowMeans(across(c(LS_1:LS_5)), na.rm = TRUE))
```

#Creating Positive and Negative Affect variable.
```{r}
qualtrics_data <- qualtrics_data %>% 
  rowwise() %>%
  mutate(
    PosAffect = rowMeans(across(c(PANAS__1, PANAS__3, PANAS__5, PANAS__9, PANAS__10, PANAS__12, 
                    PANAS__14, PANAS__16, PANAS__17, PANAS__19)), 
                      na.rm = TRUE),
    NegAffect = rowMeans(across(c(PANAS__2, PANAS__4, PANAS__6, PANAS__7, PANAS__8, PANAS__11, 
                    PANAS__13, PANAS__15, PANAS__18, PANAS__20))))
```

#Reapplying the variable and value labels that were lost from reverse scored items.
```{r}
var_label(qualtrics_data) <- var_labels
val_labels(qualtrics_data) <- val_labels
```


#Per preregistration, performing a POMP transformations on all well-being variables and
#taking their mean to create a composite Well-Being variable.
```{r}
qualtrics_data <- qualtrics_data %>%
  mutate(
    # Anxiety: scored from 1 to 4, with 4 indicating high anxiety
    Anxiety_POMP = ((Anxiety - 1) / (4 - 1)) * 100,
    # Depression: scored from 1 to 4, with 4 indicating high depression
    Depression_POMP = ((Depression - 1) / (4 - 1)) * 100,
    # Life Satisfaction: scored from 1 to 7, with 7 indicating high Life Satisfaction
    LifeSatisf_POMP = ((LifeSatisf - 1) / (7 - 1)) * 100,
    # Positive Affect: scored from 1 to 5, with 5 indicating high Positive Affect
    PosAffect_POMP = ((PosAffect - 1) / (5 - 1)) * 100,
    # Negative Affect: scored from 1 to 5, with 5 indicating high Negative Affect
    NegAffect_POMP = ((NegAffect - 1) / (5 - 1)) * 100
  ) %>%
  # Creating reversed variables for Well-Being calculation
  mutate(
    # Reverse-scoring variables where higher scores should reflect lower well-being
    Anxiety_POMP_r = 100 - Anxiety_POMP,
    Depression_POMP_r = 100 - Depression_POMP,
    NegAffect_POMP_r = 100 - NegAffect_POMP) %>% 
    # Creating the Well-Being score
    rowwise() %>%
      mutate(
        WellBeing = mean(c(Anxiety_POMP_r, Depression_POMP_r, LifeSatisf_POMP, 
                           PosAffect_POMP, NegAffect_POMP_r), na.rm = TRUE)
      ) %>%
      ungroup()
```

#Creating dataset for analyses, removing unneeded variables.
```{r}
data_for_analyses <- qualtrics_data %>% 
  select(Gender_:PANAS__20, NFM:WellBeing)
```

#Reordering variables.
```{r}
data_for_analyses <- data_for_analyses %>% 
  select(Gender_:PolLib_, NFM:WellBeing, NFM1r:PANAS__20)
```
