---
title: "Data Cleaning"
output: 
  html_document:
    theme: cosmo
    toc: true
    toc_float: true
    code_folding: show
---

## Goal

> The goal of this project is to clean a dataset for one of my thesis studies. The dataset comes directly as it was downloaded from Qualtrics. I will clean the dataset by removing cases according to the criteria from my preregisteration. I'll also combine item scores into variables and combine certain well-being variables into a composite well-being variable. The overall goal is to create a dataset that is ready for analyses.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Importing Data

```{r}
library(tidyverse)
library(labelled)
library(haven)
qualtrics_data <- read_sav("data/Data from Qualtrics - CSJA Study 2_April 1, 2025_12.34.sav")
#n = 176
#Expecting n = 150 valid cases because 150 credits were granted in SONA to respondents
#who proceeded to the end of the survey.
```

## Removing error cases

```{r}
#Per preregistration, using only cases in which participants proceeded to the end of the survey.
qualtrics_data <- qualtrics_data %>% 
  filter(Finished == 1)
#n = 159
```

```{r}
#Removing responses from SONA Admin which were used for technical troubleshooting.
qualtrics_data <- qualtrics_data %>% 
  filter(id != "9244")
#n = 157
```

## Removing duplicates

```{r}
#A technical issue in which the SONA study was not automatically granting participants credit 
#when they completed the Qualtrics survey caused some participants to take the survey more 
#than once. I'll now screen out duplicate survey responses based on the id variable.
duplicate_cases <- qualtrics_data %>%
  group_by(id) %>%
  filter(n() > 1) %>%
  arrange(id) %>% 
  dplyr::select(id, dplyr::everything())
duplicate_cases
```
There are 6 sets of duplicate responses.

```{r}
#For the first set of duplicates, the response with ResponseId = R_1YeOsLfdtzWYfHD has duration of 12
#seconds and did not proceed further than the eligibility question.
#For the second set of duplicates, both are complete responses. I will remove the response that came later,
#ResponseId = R_7n2BtvDKzUgs0xn.
#Same with the third set. Removing R_1jcEdJAa4cI8FqL.
#Same with the fourth set. Removing R_6dFnR4w70ypPgkp.
#For the fifth set, R_5390VyhkwD8FFRI did not proceed further than the eligibility question.
#For the sixth set, I will remove the later response, R_5fvLGGf4AbBIGX8.
qualtrics_data <- qualtrics_data %>%
  filter(!ResponseId %in% c(
    "R_1YeOsLfdtzWYfHD",
    "R_7n2BtvDKzUgs0xn", 
    "R_1jcEdJAa4cI8FqL",
    "R_6dFnR4w70ypPgkp",
    "R_5390VyhkwD8FFRI",
    "R_5fvLGGf4AbBIGX8"
  ))
#n = 151.
```

```{r}
#Respondent with survey id 26063 answered no the eligibility question (US citizenship)
#and was not permitted to proceed with the survey. Removing.
qualtrics_data <- qualtrics_data %>%
  filter(id != 26063)
#n = 150
```

```{r}
#Respondent with survey id 26230 did not answer any questions after the demographic 
#secion of the survey, so they have no usable data and will be removed.
qualtrics_data <- qualtrics_data %>%
  filter(id != 26230)
#n = 149
```


```{r}
#Some items were mistakenly named as reversed items. Fixing the names.
qualtrics_data <- qualtrics_data %>%
  rename(NFM8 = NFM8r, CSJAS22 = CSJAS22r)
```

## Screening out cases based on preregistration criteria

```{r}
#Per preregistration, screening for cases who selected all highest or all lowest
#response options across all surveys for exclusion from analysis.

#Calculating each participant's variance for each scale.
qualtrics_data <- qualtrics_data %>%
  mutate(
    respvar_NFM = apply(dplyr::select(., NFM1r:NFM10), 1, var, na.rm = TRUE),
    respvar_CSJAS = apply(dplyr::select(., CSJAS1:CSJAS28), 1, var, na.rm = TRUE),
    respvar_RCI = apply(dplyr::select(., RCI1:RCI10), 1, var, na.rm = TRUE),
    respvar_EXPL = apply(dplyr::select(., SEP:EL_), 1, var, na.rm = TRUE),
    respvar_ANX = apply(dplyr::select(., Anx__1:Anx__7), 1, var, na.rm = TRUE),
    respvar_DEP = apply(dplyr::select(., Dep_1:Dep_13), 1, var, na.rm = TRUE),
    respvar_LS = apply(dplyr::select(., LS_1:LS_5), 1, var, na.rm = TRUE),
    respvar_PANAS = apply(dplyr::select(., PANAS__1:PANAS__20), 1, var, na.rm = TRUE)
  )
#Identifying cases with no response variance across all scales.
qualtrics_data %>%
  filter(respvar_NFM == 0,
         respvar_CSJAS == 0,
         respvar_RCI == 0,
         respvar_EXPL == 0,
         respvar_ANX == 0,
         respvar_DEP == 0,
         respvar_LS == 0,
         respvar_PANAS == 0)
#Zero cases with no response variance across all quesionnaires. No participants excluded.
```

```{r}
#Per preregistration, screening for cases with survey completion times below 3 standard
#deviations of the mean survey completion time for exclusion from analysis.
qualtrics_data %>%
  mutate(duration_minutes = Duration__in_seconds_/60) %>% 
  mutate(
    duration_mean = mean(duration_minutes, na.rm = TRUE),
    duration_sd = sd(duration_minutes, na.rm = TRUE),
    z_duration_minutes = (duration_minutes - duration_mean)/duration_sd
  ) %>% 
  filter(z_duration_minutes <= -3)
#No participants excluded.
```

## Combining scale item scores into variable scores

```{r}
#Storing the variable labels before transformation.
var_labels <- var_label(qualtrics_data)
val_labels <- val_labels(qualtrics_data)
```

### NFM

```{r}
#Reverse scoring the appropriate NFM items.
#Examining reversed items. (Scored 1-7.)
qualtrics_data %>%
  dplyr::select(matches("^NFM.*r$")) %>%
  names()
print(qualtrics_data$NFM1r)
print(qualtrics_data$NFM3r)
print(qualtrics_data$NFM7r)
#Reversing.
qualtrics_data <- qualtrics_data %>%
  mutate(across(matches("^NFM.*r$"), ~8 - .))
#Verifying reversal.
print(qualtrics_data$NFM1r)
print(qualtrics_data$NFM3r)
print(qualtrics_data$NFM7r)

#Creating NFM variable.
qualtrics_data <- qualtrics_data %>% 
  mutate(NFM = rowMeans(across(NFM1r:NFM10), na.rm = TRUE))
```

### CSJAS

```{r}
#Reverse scoring the appropriate CSJAS items.
#Examining reversed items. (Scored 1-5.)
qualtrics_data %>%
  dplyr::select(matches("^CSJAS.*r$")) %>%
  names()
csjas_reversed <- c("CSJAS4r", "CSJAS5r", "CSJAS7r", "CSJAS11r", "CSJAS14r", 
                    "CSJAS15r", "CSJAS18r", "CSJAS20r", "CSJAS21r", "CSJAS23r",
                    "CSJAS25r", "CSJAS26r")
for(item in csjas_reversed) {
  cat("\n--- Item:", item, "---\n")
  print(qualtrics_data[[item]])
}
#Reversing.
qualtrics_data <- qualtrics_data %>%
  mutate(across(matches("^CSJAS.*r$"), ~6 - .))
#Confirming reversal.
print(qualtrics_data$CSJAS26r)

#Creating the CSJAS variable by combining the 7 final scale items determined in Study 1.
qualtrics_data <- qualtrics_data %>% 
  mutate(CSJAS = rowMeans(across(c(CSJAS1, CSJAS2, CSJAS4r, CSJAS10, CSJAS15r, CSJAS17, CSJAS28)),
                          na.rm = TRUE)) %>% 
#Rescoring to be 0-4 instead of 1-5 (to match Lahtinen).
  mutate(CSJAS = CSJAS - 1)
```

### Relig

```{r}
#Creating the Religious Commitment variable and Intrapersonal and Interpersonal subscale variables.
qualtrics_data <- qualtrics_data %>% 
  mutate(
    Relig = rowMeans(across(RCI1:RCI10), na.rm = TRUE),
    Relig_inter = rowMeans(across(c(RCI2, RCI6, RCI9, RCI10)), na.rm = TRUE),
    Relig_intra = rowMeans(across(c(RCI1, RCI3, RCI4, RCI5, RCI7, RCI8)), na.rm = TRUE)
  )
```

### WB Variables

```{r}
#Creating Anxiety variable.
qualtrics_data <- qualtrics_data %>% 
  mutate(
    Anxiety = rowMeans(across(c(Anx__1:Anx__7)), na.rm = TRUE))

#Creating Depression variable.
qualtrics_data <- qualtrics_data %>% 
  mutate(
    Depression = rowMeans(across(c(Dep_1:Dep_13)), na.rm = TRUE))

#Creating Life Satisfaction variable.
qualtrics_data <- qualtrics_data %>% 
  mutate(
    LifeSatisf = rowMeans(across(c(LS_1:LS_5)), na.rm = TRUE))

#Creating Positive and Negative Affect variable.
qualtrics_data <- qualtrics_data %>% 
  rowwise() %>%
  mutate(
    PosAffect = rowMeans(across(c(PANAS__1, PANAS__3, PANAS__5, PANAS__9, PANAS__10, PANAS__12, 
                                  PANAS__14, PANAS__16, PANAS__17, PANAS__19)), 
                         na.rm = TRUE),
    NegAffect = rowMeans(across(c(PANAS__2, PANAS__4, PANAS__6, PANAS__7, PANAS__8, PANAS__11, 
                                  PANAS__13, PANAS__15, PANAS__18, PANAS__20))))
```

### WB

```{r}
#Per preregistration, performing a POMP transformations on all well-being variables and
#taking their mean to create a composite Well-Being variable.
qualtrics_data <- qualtrics_data %>%
  mutate(
    # Anxiety: scored from 1 to 4, with 4 indicating high anxiety
    Anxiety_POMP = ((Anxiety - 1) / (4 - 1)) * 100,
    # Depression: scored from 1 to 4, with 4 indicating high depression
    Depression_POMP = ((Depression - 1) / (4 - 1)) * 100,
    # Life Satisfaction: scored from 1 to 7, with 7 indicating high Life Satisfaction
    LifeSatisf_POMP = ((LifeSatisf - 1) / (7 - 1)) * 100,
    # Positive Affect: scored from 1 to 5, with 5 indicating high Positive Affect
    PosAffect_POMP = ((PosAffect - 1) / (5 - 1)) * 100,
    # Negative Affect: scored from 1 to 5, with 5 indicating high Negative Affect
    NegAffect_POMP = ((NegAffect - 1) / (5 - 1)) * 100
  ) %>%
  # Creating reversed variables for Well-Being calculation
  mutate(
    # Reverse-scoring variables where higher scores should reflect lower well-being
    Anxiety_POMP_r = 100 - Anxiety_POMP,
    Depression_POMP_r = 100 - Depression_POMP,
    NegAffect_POMP_r = 100 - NegAffect_POMP) %>% 
  # Creating the Well-Being score
  rowwise() %>%
  mutate(
    WellBeing = mean(c(Anxiety_POMP_r, Depression_POMP_r, LifeSatisf_POMP, 
                       PosAffect_POMP, NegAffect_POMP_r), na.rm = TRUE)
  ) %>%
  ungroup()
```

## Removing WB Outlier

```{r}
#DEVIATING FROM PREREGISTRATION
# Create z-score version of WellBeing
WB_outlier <- qualtrics_data %>%
  mutate(WellBeing_z = round(abs(scale(WellBeing)[,1]), 2)) %>% 
  dplyr::select(WellBeing_z, id) %>% 
  arrange(desc(WellBeing_z))
WB_outlier %>% 
  head()
#During analyses, I discovered that this participant was dramatically skewing results.
#Excluding this case.
qualtrics_data <- qualtrics_data %>%
  filter(id != 26256)
# n = 148
```

## Dummy Coding Demographic Variables

Creating dummy variables where necessary for demographic variables which will be used as control variables in analyses.

For categorical variables, categories that comprise less than 5% of the sample will be combined.

### Gender

```{r}
#Gender
qualtrics_data %>% 
  count(Gender_)
#Gender has possible values of 1 = male, 2 = female, or 3 = Nonbinary / Third gender.
#There are no cases of Gender = 3 in the dataset, so Gender can be added to the regression 
  #analyses with no need for dummy coding.
```


### Sexual Orientation
```{r}
qualtrics_data %>% 
  count(SexOrient_) %>% 
  mutate(perc = n/1.49)
#Sexual Orientation has possible values of 1 = heterosexual, 2 = homosexual, 3 = bisexual, 
  #4 = other, and 5 = prefer not to say.
#There are 140 cases of heterosexual, 2 cases of homosexual, and 7 cases of bisexual.
#Creating dummy variable for Sexual Orientation, combining homosexual and bisexual since they 
  #make up only 1.3% and 4.7% percent of the sample, respectively.
qualtrics_data <- qualtrics_data %>%
  mutate(
    SexOrien_dummy = as.numeric(SexOrient_ %in% c(2, 3))
  )
qualtrics_data %>% 
  count(SexOrien_dummy)%>% 
  mutate(perc = n/1.49)
```

### Race
```{r}
qualtrics_data %>% 
  count(Race_) %>% 
  mutate(perc = n/1.49)
#Create dummy variables for Race categories.
#Using White (7; 75.2% of sample) as the reference category.
#Combining Black (3; 4.7%), Middle Eastern or N. African (5; .7%), and Other (8; 2.7%)
  #into Black_or_Other dummy variable.
qualtrics_data <- qualtrics_data %>%
  mutate(
    Asian_dummy = as.numeric(Race_ == 2),
    Black_or_Other_dummy = as.numeric(Race_ %in% c(3, 5, 8)),
    Hispanic_dummy = as.numeric(Race_ == 4),
  )
qualtrics_data %>% 
  count(Black_or_Other_dummy)%>% 
  mutate(perc = n/1.49)
```

### Age
```{r}
qualtrics_data <- qualtrics_data %>%
  mutate(Age = as.numeric(Age))
```

### Major
```{r}
#Values for the (academic) Major variable came from a textbox in the survey (this was a mistake). 
#I worked with Claude.ai to create this code to sort cases into major cateogories.
# This code will create two new variables: Major_Category and Major_Category2
qualtrics_data <- qualtrics_data %>%
  mutate(
    # Create Major_Category with 14 categories (including Uncategorized)
    Major_Category = case_when(
      # Category 1 - Economics
      str_detect(tolower(Major_), "^economics|^econ|^mathematical economics") ~ 1,
      
      # Category 2 - Political Science
      str_detect(tolower(Major_), "^politic|^international politics|^poly sci|^politics|^yes - politics") ~ 2,
      
      # Category 3 - Psychology
      str_detect(tolower(Major_), "^psychology|^university. psychology") |
        Major_ %in% c("double major in psychology and sociology", "Psycholgy/Communications/Marketing") ~ 3,
      
      # Category 4 - Sociology
      str_detect(tolower(Major_), "^sociology") ~ 4,
      
      # Category 5 - Business
      str_detect(tolower(Major_), "^business|^bem|^finance|^accounting|^acconting|^mathematical business") |
        Major_ %in% c("yes, business", "Business/Psych", "Finance and Computer Science", "Intended Acconting") ~ 5,
      
      # Category 6 - STEM - Biological/Medical
      str_detect(tolower(Major_), "^biology|^biochemistry|^pre-med|^pre med") |
        Major_ %in% c("medicinal chemistry", "biochemistry and molecular biology", 
                     "Biology B.S.", "Biochemistry and African-American Studies", 
                     "biology/psychology", "Pre-medicine") ~ 6,
      
      # Category 7 - Health & Exercise Science
      str_detect(tolower(Major_), "^health|^hes") ~ 7,
      
      # Category 8 - Communications
      str_detect(tolower(Major_), "^communication") ~ 8,
      
      # Category 9 - Undecided
      str_detect(tolower(Major_), "^undecided|^undeclared|^tbd|^n/a|^unsure") |
        Major_ == "Yes and my major is undecided" ~ 9,
      
      # Category 10 - STEM - Physical Sciences
      str_detect(tolower(Major_), "^chemistry|^engineering") |
        Major_ %in% c("Computer Science", "Computer Science & Finance") ~ 10,
      
      # Category 13 - Education
      str_detect(tolower(Major_), "^education|^elementary education") |
        Major_ == "Mathematics education" ~ 13,
      
      # Category 11 - STEM - Mathematics
      str_detect(tolower(Major_), "^math|^applied math") |
        Major_ == "Statistics and Psychology" ~ 11,
      
      # Category 12 - Humanities
      str_detect(tolower(Major_), "^english|^history|^theatre|^philosphy|^philosophy|^studio art") ~ 12,
      
      # Category 99 - Uncategorized
      TRUE ~ 99
    ),
    
    # Create Major_Category2 with 5 consolidated categories
    Major_Category2 = case_when(
      # Category 1 - Business & Economics
      Major_Category %in% c(1, 5) ~ 1,
      
      # Category 2 - Social Sciences
      Major_Category %in% c(2, 3, 4, 8) ~ 2,
      
      # Category 3 - STEM
      Major_Category %in% c(6, 7, 10, 11) ~ 3,
      
      # Category 4 - Humanities & Education
      Major_Category %in% c(12, 13) ~ 4,
      
      # Category 5 - Other/Undecided
      Major_Category %in% c(9, 99) ~ 5,
      
      TRUE ~ NA_real_  # Catch any unforeseen cases
    )
  )

# Add labels to the categories (optional)
qualtrics_data <- qualtrics_data %>%
  mutate(
    Major_Category_Label = case_when(
      Major_Category == 1 ~ "Economics",
      Major_Category == 2 ~ "Political Science",
      Major_Category == 3 ~ "Psychology",
      Major_Category == 4 ~ "Sociology",
      Major_Category == 5 ~ "Business",
      Major_Category == 6 ~ "STEM - Biological/Medical",
      Major_Category == 7 ~ "Health & Exercise Science",
      Major_Category == 8 ~ "Communications",
      Major_Category == 9 ~ "Undecided",
      Major_Category == 10 ~ "STEM - Physical Sciences",
      Major_Category == 11 ~ "STEM - Mathematics",
      Major_Category == 12 ~ "Humanities",
      Major_Category == 13 ~ "Education",
      Major_Category == 99 ~ "Uncategorized",
      TRUE ~ "Unknown"
    ),
    
    Major_Category2_Label = case_when(
      Major_Category2 == 1 ~ "Business & Economics",
      Major_Category2 == 2 ~ "Social Sciences",
      Major_Category2 == 3 ~ "STEM",
      Major_Category2 == 4 ~ "Humanities & Education",
      Major_Category2 == 5 ~ "Other/Undecided",
      TRUE ~ "Unknown"
    )
  )
#Examining how majors were sorted to make any adjustments.
majors<- qualtrics_data %>% 
  dplyr::select(Major_, Major_Category, Major_Category_Label, Major_Category2, Major_Category2_Label) %>% 
  arrange(Major_Category2, Major_)
#Creating dummy variables for major.
qualtrics_data %>% 
  count(Major_Category2_Label) %>% 
  mutate(perc = n/1.49)
#Using 'Business & Economics' (1; most numerous) as the reference category.
qualtrics_data <- qualtrics_data %>%
  mutate(
    SocScience_dummy = as.numeric(Major_Category2 == 2),
    STEM_dummy = as.numeric(Major_Category2 == 3),
    Hum_Educ_dummy = as.numeric(Major_Category2 == 4),
    Other_Undec_dummy = as.numeric(Major_Category2 == 5),
  )
```

```{r}
#Reapplying the variable and value labels that were lost from reverse scored items.
var_label(qualtrics_data) <- var_labels
val_labels(qualtrics_data) <- val_labels
```

```{r}
#Renaming some variables.
qualtrics_data <- qualtrics_data %>%
  rename(Gender = Gender_, SexOrien = SexOrient_, Race = Race_,
         Major = Major_, PolLib = PolLib_)
```

## Creating Dataset for Analysis

```{r}
#Creating dataset for analysis and removing unneeded variables.
data_for_analysis <- qualtrics_data %>% 
  dplyr::select(Gender:PANAS__20, NFM:Other_Undec_dummy)
```

```{r}
#Reordering variables and writing out dataset.
data_for_analysis <- data_for_analysis %>% 
  dplyr::select(Gender:Age, Major_Category2_Label, PolLib, NFM:WellBeing, NFM1r:PANAS__20,
         Major, Major_Category, Major_Category2, Major_Category_Label, 
         SexOrien_dummy:Hispanic_dummy, SocScience_dummy:Other_Undec_dummy)
write_sav(data_for_analysis, "data/data_for_analysis.sav")
```
