---
title: "Portfolio 6"
output: 
  html_document:
    theme: cosmo
    toc: true
    toc_float: true
    code_folding: show
---

> The projects should be numbered consecutively (i.e., in the order in which you began them), and should include for each project a description of the goal, the product (computer program, hand graph, computer graph, etc.), the data, and some interpretation. Reports must be reproducible and of high quality in terms of writing, grammar, presentation, etc.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = FALSE)
```

```{r eval = FALSE}
Relig_vars <- c("scale(Relig)", "scale(Relig_inter)", "scale(Relig_intra)")
expl_vars <- c("scale(NFM)")

# Function to run regressions using formulas created from strings
run_regressions <- function(dependent_var, explanatory_var) {
  # Model 1: Just the explanatory variable
  print(paste("Model 1:", dependent_var, "~", explanatory_var))
  print(summary(lm(as.formula(paste(dependent_var, "~",
                                    explanatory_var)), data = 
                     data_for_analysis)))
  
  # Model 2: Add all controls
  print(paste("Model 2:", dependent_var, "~", explanatory_var, "+ all controls"))
  print(summary(lm(as.formula(paste(dependent_var, "~", 
                                    explanatory_var, 
    "+ PolLib + Gender + SexOrien_dummy + Asian_dummy +
    Black_or_Other_dummy + Hispanic_dummy + SocScience_dummy + 
    STEM_dummy + Hum_Educ_dummy + Other_Undec_dummy")), 
    data = data_for_analysis)))
}

# Run all combinations of dependent and explanatory variables
for(i in 1:length(Relig_vars)) {
  for(j in 1:length(expl_vars)) {
    run_regressions(Relig_vars[i], expl_vars[j])
  }
}
```

```{r eval = FALSE}
record_regr_output <- function()
regr_summary <- summary(lm(scale(Relig) ~ scale(NFM), data = data_for_analysis))

predictor <- regr_summary$terms[[3]][[2]]
outcome <- regr_summary$terms[[2]][[2]]
beta <- regr_summary$coefficients[2]
t <- regr_summary$coefficients[6]
t_p <- regr_summary$coefficients[8]
R <- regr_summary$r.squared
adjR <- regr_summary$adj.r.squared
F_ <- print(paste0("F(", regr_summary$fstatistic[2], ",", regr_summary$fstatistic[3], ") = ", round(regr_summary$fstatistic[1], 2))) #Claude helped improve this by removing the extra spaces and rounding.
F_p <- pf(regr_summary$fstatistic[1], regr_summary$fstatistic[2], regr_summary$fstatistic[3], lower.tail = FALSE) #Claude gave me this code.

regr_output <- tibble(
  model = model,
  beta = beta,
  t = t,
  t_p = t_p,
  R = R,
  adjR = adjR,
  F_ = F_,
  F_p = F_p
)

regr_output
```

I couldn't figure out how to combine these two pieces, so I asked Claude for help.

```{r}
Relig_vars <- c("scale(Relig)", "scale(Relig_inter)", "scale(Relig_intra)")
expl_vars <- c("scale(NFM)")

# Initialize an empty tibble to store all results
all_results <- tibble()

# Function to run regressions and store results
run_regressions <- function(dependent_var, explanatory_var) {
  # Model 1: Just the explanatory variable
  model_name <- paste("Model 1:", dependent_var, "~", explanatory_var)
  regr_summary <- summary(lm(as.formula(paste(dependent_var, "~",
                                   explanatory_var)), data = data_for_analysis))
  
  # Store results
  beta <- regr_summary$coefficients[2, 1]
  t <- regr_summary$coefficients[2, 3]
  t_p <- regr_summary$coefficients[2, 4]
  R <- regr_summary$r.squared
  adjR <- regr_summary$adj.r.squared
  F_ <- paste0("F(", regr_summary$fstatistic[2], ",", regr_summary$fstatistic[3], ") = ", round(regr_summary$fstatistic[1], 2))
  F_p <- pf(regr_summary$fstatistic[1], regr_summary$fstatistic[2], regr_summary$fstatistic[3], lower.tail = FALSE)
  
  # Create row and bind to results
  regr_output <- tibble(
    model = model_name,
    predictor = explanatory_var,
    outcome = dependent_var,
    beta = beta,
    t = t,
    t_p = t_p,
    R = R,
    adjR = adjR,
    F_ = F_,
    F_p = F_p
  )
  
  all_results <<- bind_rows(all_results, regr_output)
  
  # Model 2: Add all controls
  model_name <- paste("Model 2:", dependent_var, "~", explanatory_var, "+ controls")
  regr_summary <- summary(lm(as.formula(paste(dependent_var, "~", 
                                   explanatory_var, 
                                   "+ PolLib + Gender + SexOrien_dummy + Asian_dummy +
                                   Black_or_Other_dummy + Hispanic_dummy + SocScience_dummy + 
                                   STEM_dummy + Hum_Educ_dummy + Other_Undec_dummy")), 
                              data = data_for_analysis))
  
  # Store results for model 2
  beta <- regr_summary$coefficients[2, 1]
  t <- regr_summary$coefficients[2, 3]
  t_p <- regr_summary$coefficients[2, 4]
  R <- regr_summary$r.squared
  adjR <- regr_summary$adj.r.squared
  F_ <- paste0("F(", regr_summary$fstatistic[2], ",", regr_summary$fstatistic[3], ") = ", round(regr_summary$fstatistic[1], 2))
  F_p <- pf(regr_summary$fstatistic[1], regr_summary$fstatistic[2], regr_summary$fstatistic[3], lower.tail = FALSE)
  
  regr_output <- tibble(
    model = model_name,
    predictor = explanatory_var,
    outcome = dependent_var,
    beta = beta,
    t = t,
    t_p = t_p,
    R = R,
    adjR = adjR,
    F_ = F_,
    F_p = F_p
  )
  
  all_results <<- bind_rows(all_results, regr_output)
}

# Run all combinations of dependent and explanatory variables
for(i in 1:length(Relig_vars)) {
  for(j in 1:length(expl_vars)) {
    run_regressions(Relig_vars[i], expl_vars[j])
  }
}

# View the results table
all_results
```