---
title: "Portfolio 7 - Cross-Referencing and Combining Datasets"
output: 
  html_document:
    theme: cosmo
    toc: true
    toc_float: true
    code_folding: show
---


## Goal

> Note: This project invovled working with data with personal identifiers; however, no personal identifiers are visisble or accessible in this project.

> Goal: As part of my TA duties, I was asked to create a spreadsheet that contains the names, emails, classification, expected completion date, and induction date of WFU seniors who are anticipated to graduate in 2025 who are current members of Psi Chi. However, all this information was not located in one place. I had a spreadsheet of all current Psi Chi WFU chapter members since the chapter began (which contains names and induction date), many of whom are not current WFU seniors expecting to graduate nor current WFU students at all. And I had another spreadsheet of all WFU psychology majors (which contains names, emails, classification, and expected completion date). Rather than manaully trying to cross reference these two spreadsheets and compile the relevant information, I decided to use R to combine the datasets using certain criteria. The final product will be a spreadsheet that contains all WFU senior Psi Chi members.

```{r setup, include=TRUE}
knitr::opts_chunk$set(echo = TRUE, eval=TRUE)
```

## Importing Data

```{r message=FALSE}
library(tidyverse)
library(stringr)
library(readr)
library(readxl)
psichi_members <- read_csv("data/Psi Chi member list March 2025.xls.xlsx - Psi Chi member list March 2025.csv")
psych_majors <- read_excel("data/WFU_-_OUR_-_Burst_Student_Information_Profile_-_by_Major_3.6.2025 Psi Chi.xlsx")
new_members <- read_csv("data/new_members.csv")
```

## Matching Formatting of Names

Want to make sure there is a column in both spreadsheets with names formatted the same way.
For some reason, in one spreadsheet, some names had "(Private)" at the beginning. Also, in one spreadsheet names were formatted "LastName, FirstName" but were formmated "FirstName LastName" in the other.

```{r}
psych_majors <- psych_majors %>%
  mutate(
    # Remove "(Private)" suffix if present
    name = str_remove(Student, " \\(Private\\)"),
    # Names are already in "Lastname, Firstname" format, so just keep them as is
    # but ensure they're clean
    name = str_trim(name)
  )

psichi_members <- psichi_members %>%
  mutate(
    # Convert "Lastname Firstname" to "Lastname, Firstname"
    # First check if it already has a comma
    name = case_when(
      str_detect(`Applicant Name`, ",") ~ `Applicant Name`,  # Already in "Lastname, Firstname" format
      TRUE ~ str_replace(`Applicant Name`, "(\\S+)\\s+(\\S+.*)", "\\1, \\2")  # Convert "Lastname Firstname" to "Lastname, Firstname"
    ),
    # Trim any extra spaces
    name = str_trim(name)
  )
```

## Joining Datasets by Name

```{r}
# Join the datasets to find psychology majors who are Psi Chi members
current_student_psichi_members <- inner_join(
  psych_majors,
  psichi_members,
  by = "name"
) %>% 
  select(name, `WFU Email Address`, `Class Standing at Start of Period`, `Program of Study Completion Status`, `Expected Completion Date`, `Induction Date`)

current_student_psichi_members <- current_student_psichi_members %>%
  mutate(`Expected Completion Date` = as.character(`Expected Completion Date`),
         email = "") %>%  # Create empty email column
  relocate(email, .after = name) %>% 
  mutate(`In Psych Majors Spreadsheet` = TRUE)
```

## Adding New Members

I also have a spreadsheet which contains Psi Chi members who should appear on the final spreadsheet but are not located in the current Psi Chi members spreadsheet.

Here I create a spreadsheet which contains the students who are new Psi Chi members and who are also in the Psych Majors spreadsheet. I also had to make sure that students who were on the new Psi Chi members spreadsheet but who were not on the Psych Majors spreadsheet were also added (because some new Psi Chi memebers were sophomors who had not yet officially declared their major). I document this with a `In Psych Majors Spreadsheet` variable. This results in some rows (students) having missing data on some variables which were only contained in the Psych Majors spreadsheet.

```{r}
current_student_psichi_new_members <- left_join(
  new_members,
  psych_majors,
  by = "name"
) %>% 
  mutate(`In Psych Majors Spreadsheet` = !is.na(`WFU Email Address`), # Create indicator variable
         `Induction Date` = "4/22/2025") %>% 
  select(name, email, `WFU Email Address`, `Class Standing at Start of Period`, `Program of Study Completion Status`, `Expected Completion Date`, `Induction Date`, `In Psych Majors Spreadsheet`)

current_student_psichi_new_members <- current_student_psichi_new_members %>%
  mutate(`Expected Completion Date` = as.character(`Expected Completion Date`),
         `Program of Study Completion Status` = ifelse(`In Psych Majors Spreadsheet` == FALSE, 
                                                      "Unknown", 
                                                      `Program of Study Completion Status`))
```

## Binding Rows to Create Final Dataset

```{r}
current_student_psichi_members <- bind_rows(current_student_psichi_members, current_student_psichi_new_members)
```

Here is a glimpse at what the final dataset looks like (excluding personal identifiers). It actually contains all Psi Chi members who are current WFU students and contains variables from which graduating seniors can be identified (which was the goal of this project).

```{r}
current_student_psichi_members %>% 
  select(-name, -`WFU Email Address`, -email)
```

Here is the code where I exported the dataset as an excel file.

```{r eval=FALSE}
write_excel_csv(current_student_psichi_members, file = "/Users/noahbooker/Documents/Documents/TA/Office/Psi Chi/current_student_psichi_members_Spring2025.csv")
```

