---
title: "Portfolio 8 - Main Analyses - Study 2"
output: 
  html_document:
    theme: cosmo
    toc: true
    toc_float: true
    code_folding: show
---
## Goal

> The goal of this project is to demonstrate how the products projects 5 and 6 


> The projects should be numbered consecutively (i.e., in the order in which you began them), and should include for each project a description of the goal, the product (computer program, hand graph, computer graph, etc.), the data, and some interpretation. Reports must be reproducible and of high quality in terms of writing, grammar, presentation, etc.


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r message = FALSE, echo=TRUE, eval=TRUE}
library(tidyverse)
library(lm.beta)
library(haven)
data_for_analysis <- read_sav("data/data_for_analysis.sav")
```

# Simplest Version of Main Analyses

### 1. Hypothesis: NFM will be positively associated with CSJAS.

Analysis: Regression of Religiosity on NFM.

Relig ~ NFM

```{r echo=TRUE}
summary(lm.beta(lm(CSJAS ~ NFM, data = data_for_analysis)))
```

### 2. Hypothesis: NFM will be positively associated with Religiosity.

Analysis: Regression of CSJAS on NFM.

CSJAS ~ NFM

```{r echo=TRUE}
summary(lm.beta(lm(Relig ~ NFM, data = data_for_analysis)))
```

### 3. Hypothesis: CSJA and Religiosity will be negatively correlated with one another.

Analyses: Pearson correlation of CSJAS and Religiosity

r(CSJAS, Relig)

```{r echo=TRUE}
data_for_analysis %>%
  with(cor.test(CSJAS, Relig, method = "pearson"))
```

### 4. Hypothesis: CSJA will negatively moderate the positive association between NFM and Religiosity.

Analysis: Regression of Religiosity on NFM, CSJA, and CSJAxNFM.

Relig ~ NFM + CSJAS + CSJASxNFM

```{r echo=TRUE}
#Creating centered variables.
data_for_analysis <- data_for_analysis %>% 
  mutate(NFM_cent = NFM - mean(NFM, na.rm = TRUE),
         CSJAS_cent = CSJAS - mean(CSJAS, na.rm = TRUE))
#Running regression.
summary(lm.beta(lm(Relig ~ CSJAS_cent*NFM_cent, data = data_for_analysis)))
```

### 5. Hypothesis: Religiosity will negatively moderate the positive relationship between NFM and CSJA.

Analysis: Regression of CSJA on NFM, Religiosity, and Religiosity*NFM.

CSJAS ~ NFM + Relig + ReligxNFM

```{r echo=TRUE}
data_for_analysis <- data_for_analysis %>% 
  mutate(Relig_cent = Relig - mean(Relig, na.rm = TRUE),
         Relig_inter_cent = Relig_inter - mean(Relig_inter, na.rm = TRUE),  #For later
         Relig_intra_cent = Relig_intra - mean(Relig_intra, na.rm = TRUE))
summary(lm.beta(lm(CSJAS ~ Relig_cent*NFM_cent, data = data_for_analysis)))
```

### 6. Hypothesis: Religiosity will be positively associated with Well-Being.

Analysis: Regression of Well-Being on Religiosity.

WB ~ Relig

```{r echo=TRUE}
summary(lm.beta(lm(WellBeing ~ Relig, data = data_for_analysis)))
```

### 7. Hypothesis: CSJA will be negatively associated with Well-Being.

Analysis: Regression of Well-Being on CSJAS.

WB ~ CSJAS

```{r echo=TRUE}
summary(lm.beta(lm(WellBeing ~ CSJAS, data = data_for_analysis)))
```

# Analyses with control variables, religiosity subscales, and constituent well-being variables

```{r}
# Function to run regressions and return formatted results
compile_regressions <- function(outcome_var, explanatory_var, dataset) {
  # Results for both models
  results <- tibble()
  
  # Model 1: No controls
  regr_summary1 <- summary(lm.beta(lm(as.formula(paste(outcome_var, "~",
                                   explanatory_var)), data = dataset)))
  
  # Create first row
  results <- tibble(
    model = "no controls",
    outcome = outcome_var,
    predictor = explanatory_var,
    b = sprintf("%.2f", regr_summary1$coefficients[2, 1]),
    B = sprintf("%.2f", regr_summary1$coefficients[2, 2]),
    SE = sprintf("%.2f", regr_summary1$coefficients[2, 3]),
    t = sprintf("%.2f", regr_summary1$coefficients[2, 4]),
    t_p = sprintf("%.3f", regr_summary1$coefficients[2, 5]),
    R = sprintf("%.2f", regr_summary1$r.squared),
    adjR = sprintf("%.2f", regr_summary1$adj.r.squared),
    F_ = round(regr_summary1$fstatistic[1], 2),
    df = paste0("(", regr_summary1$fstatistic[2], ", ", regr_summary1$fstatistic[3], ")"),
    F_p = sprintf("%.3f", pf(regr_summary1$fstatistic[1], regr_summary1$fstatistic[2], 
                           regr_summary1$fstatistic[3], lower.tail = FALSE))
  )
  
  # Model 2: With controls
  regr_summary2 <- summary(lm.beta(lm(as.formula(paste(outcome_var, "~", 
                                   explanatory_var, 
                                   "+ PolLib + Gender + Age + SexOrien_dummy + Asian_dummy +
                                   Black_or_Other_dummy + Hispanic_dummy + SocScience_dummy + 
                                   STEM_dummy + Hum_Educ_dummy + Other_Undec_dummy")), 
                              data = dataset)))
  
  # Add second row
  results <- bind_rows(results, tibble(
    model = "w/ controls",
    outcome = outcome_var,
    predictor = explanatory_var,
    b = sprintf("%.2f", regr_summary2$coefficients[2, 1]),
    B = sprintf("%.2f", regr_summary2$coefficients[2, 2]),
    SE = sprintf("%.2f", regr_summary2$coefficients[2, 3]),
    t = sprintf("%.2f", regr_summary2$coefficients[2, 4]),
    t_p = sprintf("%.3f", regr_summary2$coefficients[2, 5]),
    R = sprintf("%.2f", regr_summary2$r.squared),
    adjR = sprintf("%.2f", regr_summary2$adj.r.squared),
    F_ = round(regr_summary2$fstatistic[1], 2),
    df = paste0("(", regr_summary2$fstatistic[2], ", ", regr_summary1$fstatistic[3], ")"),
    F_p = sprintf("%.3f", pf(regr_summary2$fstatistic[1], regr_summary2$fstatistic[2], 
                           regr_summary2$fstatistic[3], lower.tail = FALSE))
  ))
  
  return(results)
}

#Function to loop the compile_regression function for combinations of variables.
loop_compile_regressions <- function(outcome_var, explanatory_var, dataset) {
  results <- tibble()
  for(i in 1:length(outcome_vars)) {
    for(j in 1:length(expl_vars)) {
      results <- bind_rows(results, 
                               compile_regressions(outcome_var[i], explanatory_var[j], dataset))
      }
  }
    return(results)
}
```

### 1. CSJAS ~ NFM

Control variables: political orientation, gender, age, race, sexual orientation, and academic major.

```{r}
outcome_vars <- c("CSJAS")
expl_vars <- c("NFM")

results2 <- loop_compile_regressions(outcome_vars, expl_vars, data_for_analysis)

options(width = 2000)           # Increase console width
options(tibble.width = Inf)     # Show all columns
options(pillar.min_chars = Inf) # Don't truncate character columns
# View results
results2
```

### 2. Relig ~ NFM 

Control variables: political orientation, gender, age, race, sexual orientation, and academic major.

```{r}
outcome_vars <- c("Relig", "Relig_inter", "Relig_intra")
expl_vars <- "NFM"

results1 <- loop_compile_regressions(outcome_vars, expl_vars, data_for_analysis)

options(width = 2000)           # Increase console width
options(tibble.width = Inf)     # Show all columns
options(pillar.min_chars = Inf) # Don't truncate character columns
# View results
results1
```


```{r message = FALSE}
library(ppcor)
```

```{r}
# Function to run correlation using strings
run_corr <- function(variable1, variable2, dataset) {
  # Simple correlation
  corr <- paste0("r(", variable1, ", ", variable2, ")")
  corr_results <- cor.test(dataset[[variable1]], dataset[[variable2]], method = "pearson")
  
  corr_output <- tibble(
    corr = corr,
    r = sprintf("%.2f", corr_results$estimate),
    p = sprintf("%.3f", corr_results$p.value),
    t = sprintf("%.2f", corr_results$statistic),
    df = corr_results$parameter
  )
  
  # 2. Partial correlation with controls
  corr <- paste0("r(", variable1, ", ", variable2, " | controls)")
  corr_results <- pcor.test(x = dataset[[variable1]], 
                    y = dataset[[variable2]], 
                    z = dataset[, c("PolLib", "Gender", "Age", "Asian_dummy", 
                                              "Black_or_Other_dummy", "Hispanic_dummy", 
                                              "SexOrien_dummy", "SocScience_dummy", "STEM_dummy", 
                                              "Hum_Educ_dummy", "Other_Undec_dummy")],
                    method = "pearson")
  
  corr_output <- bind_rows(corr_output, tibble(
    corr = corr,
    r = sprintf("%.2f", corr_results$estimate),
    p = sprintf("%.3f", corr_results$p.value),
    t = sprintf("%.2f", corr_results$statistic),
    df = corr_results$n - corr_results$gp - 2
  ))
  return(corr_output)
}
# Run all combinations of variables
loop_corr <- function(variable1, variable2, dataset) {
  results <- tibble()
  for(i in 1:length(variable1)) {
    for(j in 1:length(variable2)) {
      results <- bind_rows(results, run_corr(variable1[i], variable2[j], dataset))
      }
    }
  return(results)
}
```

### 3. r(CSJAS, Relig)

Control variables: political orientation, gender, age, race, sexual orientation, and academic major.

```{r}
var1 <- c("CSJAS")
vars2 <- c("Relig", "Relig_inter", "Relig_intra")

results3 <- loop_corr(var1, vars2, data_for_analysis)

options(width = 2000)           # Increase console width
options(tibble.width = Inf)     # Show all columns
options(pillar.min_chars = Inf) # Don't truncate character columns
# View the results table
results3
```

```{r}
compile_mod_regressions <- function(outcome_var, explanatory_var, dataset) {
  # Results for both models
  results <- tibble()
  
  # Model 1: No controls
  regr_summary1 <- summary(lm.beta(lm(as.formula(paste(outcome_var, "~",
                                   explanatory_var)), data = dataset)))
  
  # Create first row
  results <- tibble(
    model = "no controls",
    outcome = outcome_var,
    predictor = explanatory_var,
    b = sprintf("%.2f", regr_summary1$coefficients[4, 1]),
    B = sprintf("%.2f", regr_summary1$coefficients[4, 2]),
    SE = sprintf("%.2f", regr_summary1$coefficients[4, 3]),
    t = sprintf("%.2f", regr_summary1$coefficients[4, 4]),
    t_p = sprintf("%.3f", regr_summary1$coefficients[4, 5]),
    R = sprintf("%.2f", regr_summary1$r.squared),
    adjR = sprintf("%.2f", regr_summary1$adj.r.squared),
    F_ = round(regr_summary1$fstatistic[1], 2),
    df = paste0("(", regr_summary1$fstatistic[2], ", ", regr_summary1$fstatistic[3], ")"),
    F_p = sprintf("%.3f", pf(regr_summary1$fstatistic[1], regr_summary1$fstatistic[2], 
                           regr_summary1$fstatistic[3], lower.tail = FALSE))
  )
 
  # Model 2: With controls
  regr_summary2 <- summary(lm.beta(lm(as.formula(paste(outcome_var, "~", 
                                   explanatory_var, 
                                   "+ PolLib + Gender + Age + SexOrien_dummy + Asian_dummy +
                                   Black_or_Other_dummy + Hispanic_dummy + SocScience_dummy + 
                                   STEM_dummy + Hum_Educ_dummy + Other_Undec_dummy")), 
                              data = dataset)))
  
  # Add second row
  results <- bind_rows(results, tibble(
    model = "w/ controls",
    outcome = outcome_var,
    predictor = explanatory_var,
    b = sprintf("%.2f", regr_summary2$coefficients[15, 1]),
    B = sprintf("%.2f", regr_summary2$coefficients[15, 2]),
    SE = sprintf("%.2f", regr_summary2$coefficients[15, 3]),
    t = sprintf("%.2f", regr_summary2$coefficients[15, 4]),
    t_p = sprintf("%.3f", regr_summary2$coefficients[15, 5]),
    R = sprintf("%.2f", regr_summary2$r.squared),
    adjR = sprintf("%.2f", regr_summary2$adj.r.squared),
    F_ = round(regr_summary2$fstatistic[1], 2),
    df = paste0("(", regr_summary2$fstatistic[2], ", ", regr_summary1$fstatistic[3], ")"),
    F_p = sprintf("%.3f", pf(regr_summary2$fstatistic[1], regr_summary2$fstatistic[2], 
                           regr_summary2$fstatistic[3], lower.tail = FALSE))
  ))
  
  return(results)
}

#Function to loop the compile_regression function for combinations of variables.
loop_compile_mod_regressions <- function(outcome_var, explanatory_var, dataset) {
  results <- tibble()
  for(i in 1:length(outcome_var)) {
    for(j in 1:length(explanatory_var)) {
      results <- bind_rows(results, 
                               compile_mod_regressions(outcome_var[i], explanatory_var[j], dataset))
      }
  }
    return(results)
}
```

### 4. Relig ~ NFM + CSJAS + CSJASxNFM

Control variables: political orientation, gender, age, race, sexual orientation, and academic major.

```{r}
outcome_vars <- c("Relig", "Relig_inter", "Relig_intra")
expl_vars <- c("CSJAS_cent*NFM_cent")

results4 <- loop_compile_mod_regressions(outcome_vars, expl_vars, data_for_analysis)

options(width = 2000)           # Increase console width
options(tibble.width = Inf)     # Show all columns
options(pillar.min_chars = Inf) # Don't truncate character columns
# View results
results4
```

### 5. CSJAS ~ NFM + Relig + ReligxNFM

Control variables: political orientation, gender, age, race, sexual orientation, and academic major.

```{r}
outcome_vars <- c("CSJAS")
expl_vars <- c("Relig_cent*NFM_cent", "Relig_inter_cent*NFM_cent", "Relig_intra_cent*NFM_cent")

results5 <- loop_compile_mod_regressions(outcome_vars, expl_vars, data_for_analysis)

options(width = 2000)           # Increase console width
options(tibble.width = Inf)     # Show all columns
options(pillar.min_chars = Inf) # Don't truncate character columns
# View results
results5
```

### 6. WB ~ Relig

Control variables: political orientation, gender, age, race, sexual orientation, and academic major.
```{r}
outcome_vars <- c("WellBeing", "Anxiety_POMP", "Depression_POMP", "LifeSatisf_POMP", 
                 "PosAffect_POMP", "NegAffect_POMP")
expl_vars <- c("Relig", "Relig_inter", "Relig_intra")

results6 <- loop_compile_regressions(outcome_vars, expl_vars, data_for_analysis)

options(width = 2000)           # Increase console width
options(tibble.width = Inf)     # Show all columns
options(pillar.min_chars = Inf) # Don't truncate character columns

print(results6, n = 36)
```


```{r eval = FALSE, echo=FALSE}
# Define variable names
outcome_vars <- c("WellBeing", "Anxiety_POMP", "Depression_POMP", "LifeSatisf_POMP", 
             "PosAffect_POMP", "NegAffect_POMP")
expl_vars <- c("Relig", "Relig_inter", "Relig_intra")

# Function to run regressions using formulas created from strings
run_regressions <- function(outcome_var, explanatory_var, dataset) {
  # Model 1: Just the explanatory variable
  print(paste("Model 1:", outcome_var, "~", explanatory_var))
  print(summary(lm.beta(lm(as.formula(paste(outcome_var, "~", explanatory_var)), data = dataset)))
  
  # Model 2: Add PolLib
  print(paste("Model 2:", outcome_var, "~", explanatory_var, "+ PolLib"))
  print(summary(lm.beta(lm(as.formula(paste(outcome_var, "~", explanatory_var, "+ PolLib")), data = dataset)))
  
  # Model 3: Add Gender
  print(paste("Model 3:", outcome_var, "~", explanatory_var, "+ PolLib + Gender"))
  print(summary(lm.beta(lm(as.formula(paste(outcome_var, "~", explanatory_var, "+ PolLib + Gender")), data = dataset)))
  
  # Model 4: Add all controls
  print(paste("Model 4:", outcome_var, "~", explanatory_var, "+ all controls"))
  print(summary(lm.beta(lm(as.formula(paste(outcome_var, "~", explanatory_var, 
    "+ PolLib + Gender + Age + SexOrien_dummy + Asian_dummy + Black_or_Other_dummy +
    Hispanic_dummy + SocScience_dummy + STEM_dummy + Hum_Educ_dummy + 
    Other_Undec_dummy")), data = dataset)))
}

# Run all combinations of outcome and explanatory variables
for(i in 1:length(outcome_vars)) {
  for(j in 1:length(expl_vars)) {
    run_regressions(outcome_vars[i], expl_vars[j])
  }
}
```

### 7. WB ~ CSJAS

Control variables: political orientation, gender, age, race, sexual orientation, and academic major.
```{r}
# Define variable names
outcome_vars <- c("WellBeing", "Anxiety_POMP", "Depression_POMP", "LifeSatisf_POMP", 
                 "PosAffect_POMP", "NegAffect_POMP")
expl_vars <- c("CSJAS")

results7 <- loop_compile_regressions(outcome_vars, expl_vars, data_for_analysis)

options(width = 2000)           # Increase console width
options(tibble.width = Inf)     # Show all columns
options(pillar.min_chars = Inf) # Don't truncate character columns

# View results
results7
```

# Full output for all analyses

x

### 1. CSJAS ~ NFM

```{r eval = TRUE, echo = FALSE}
outcome_vars <- c("CSJAS")
expl_vars <- c("NFM")

# Run all combinations of outcome and explanatory variables
for(i in 1:length(outcome_vars)) {
  for(j in 1:length(expl_vars)) {
    run_regressions(outcome_vars[i], expl_vars[j], data_for_analysis)
  }
}
```

### 2. Relig ~ NFM

```{r eval = TRUE, echo = FALSE}
outcome_vars <- c("Relig", "Relig_inter", "Relig_intra")
expl_vars <- c("NFM")

# Run all combinations of outcome and explanatory variables
for(i in 1:length(outcome_vars)) {
  for(j in 1:length(expl_vars)) {
    run_regressions(outcome_vars[i], expl_vars[j], data_for_analysis)
  }
}
```

### 3. r(CSJAS, Relig)

```{r echo = FALSE}
var1 <- c("CSJAS")
vars2 <- c("Relig", "Relig_inter", "Relig_intra")

# Function to run correlation using strings
run_corr <- function(variable1, variable2) {
  # Simple correlation
  print(paste0("r(", variable1, ", ", variable2, ")"))
  print(cor.test(data_for_analysis[[variable1]], data_for_analysis[[variable2]], method = "pearson"))
  
  # Partial correlation with controls
  print(paste0("r(", variable1, ", ", variable2, "|controls)"))
  result <- pcor.test(x = data_for_analysis[[variable1]], 
                    y = data_for_analysis[[variable2]], 
                    z = data_for_analysis[, c("PolLib", "Gender", "Age", "Asian_dummy", 
                                              "Black_or_Other_dummy", "Hispanic_dummy", 
                                              "SexOrien_dummy", "SocScience_dummy", "STEM_dummy", 
                                              "Hum_Educ_dummy", "Other_Undec_dummy")],
                    method = "pearson")
  print(result)
  cat("\n-------------------\n\n")
}

# Run all combinations of variables
for(i in 1:length(var1)) {
  for(j in 1:length(vars2)) {
    run_corr(var1[i], vars2[j])
  }
}
```

### 4. Relig ~ NFM + CSJAS + CSJASxNFM

```{r}

outcome_vars <- c("Relig", "Relig_inter", "Relig_intra")


run_mod_regressions <- function(outcome_var, explanatory_var, dataset) {
  # Model 1: Just NFM and CSJAS with interaction
  print(paste("Model 1:", outcome_var, "~", explanatory_var))
  print(summary(lm.beta(lm(as.formula(paste(outcome_var, "~", explanatory_var)), 
                   data = dataset))))
  
  # Model 2: Add all controls
  print(paste("Model 2:", outcome_var, "~", explanatory_var, "+ all controls"))
  print(summary(lm.beta(lm(as.formula(paste(outcome_var, "~", explanatory_var,
    "+ PolLib + Gender + Age + SexOrien_dummy + Asian_dummy +
    Black_or_Other_dummy + Hispanic_dummy + SocScience_dummy + 
    STEM_dummy + Hum_Educ_dummy + Other_Undec_dummy")), 
    data = dataset))))
  cat("\n-------------------\n\n")
}


loop_run_mod_regressions <- function(outcome_var, explanatory_var, dataset) {
  for(i in 1:length(outcome_var)) {
    for(j in 1:length(explanatory_var)) {
      run_mod_regressions(outcome_var[i], explanatory_var[j], dataset)
    }}}

outcome_vars <- c("Relig", "Relig_inter", "Relig_intra")
expl_vars <- c("CSJAS_cent*NFM_cent")

loop_run_mod_regressions(outcome_vars, expl_vars, data_for_analysis)
```

### 5. CSJAS ~ NFM + Relig + ReligxNFM

```{r}
outcome_vars <- c("CSJAS")
expl_vars <- c("Relig_cent*NFM_cent", "Relig_inter_cent*NFM_cent", "Relig_intra_cent*NFM_cent")

loop_run_mod_regressions(outcome_vars, expl_vars, data_for_analysis)
```

### 6. WB ~ Relig

```{r}
# Define variable names
outcome_vars <- c("WellBeing", "Anxiety_POMP", "Depression_POMP", "LifeSatisf_POMP", 
             "PosAffect_POMP", "NegAffect_POMP")
expl_vars <- c("Relig", "Relig_inter", "Relig_intra")

# Run all combinations of outcome and explanatory variables
for(i in 1:length(outcome_vars)) {
  for(j in 1:length(expl_vars)) {
    run_regressions(outcome_vars[i], expl_vars[j], data_for_analysis)
  }
}
```

### 7. WB ~ CSJAS

```{r}
# Define variable names
outcome_vars <- c("WellBeing", "Anxiety_POMP", "Depression_POMP", "LifeSatisf_POMP", 
             "PosAffect_POMP", "NegAffect_POMP")
expl_vars <- c("CSJAS")

# Run all combinations of outcome and explanatory variables
for(i in 1:length(outcome_vars)) {
  for(j in 1:length(expl_vars)) {
    run_regressions(outcome_vars[i], expl_vars[j], data_for_analysis)
  }
}
```

