---
title: "Portfolio 9 - Confirmatory Factor Analysis"
output: 
  html_document:
    theme: cosmo
    toc: true
    toc_float: true
    code_folding: show
---

## Goal

> The goal of this project is to run a confirmatory factor analysis (CFA) on the a seven-item critical social justice attitudes scale (CSJAS) which was previosuly created using exploratory factor analysis (EFA) in SPSS. I also run a CFA on another seven-item scale used in the study, the GAD-7 anxiety scale, to compare the fit indices.

> The data is from 149 undergraduates. The CFA is run on a randomly split half of the sample.

> The final product is the interpretation of the CFA which was used in the write-up of this study.


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Loading data and packages

```{r}
library(haven)
DataForAnalysis <- read_sav("data/DataForAnalysis.sav")
library(rsample)
library(lavaan)
```

## Setting seed and splitting data

Note: Claude.ai assisted in creating the code for this project.

```{r}
# Set a seed for reproducibility
set.seed(1234)

# Create an initial split with 50/50 proportion
data_split <- initial_split(DataForAnalysis, prop = 0.5)

# Save training data
EFA_data <- training(data_split)
dim(EFA_data)

# Save testing data
CFA_data <- testing(data_split)
dim(CFA_data)
```


```{r}
#Write out EFA data to replicate EFA in SPSS.
#write_sav(EFA_data, "EFA_data.sav")
```

## Conducting CFA

```{r}
# Define your CFA model with a different name for the latent factor
cfa_model <- '
  # Define one-factor structure with a different name
  CSJAS_Factor =~ CSJAS1 + CSJAS2 + CSJAS4r + CSJAS10 + CSJAS15r + CSJAS17 + CSJAS28
'

# Fit the CFA model using maximum likelihood estimation
cfa_fit <- cfa(cfa_model, 
               data = CFA_data, 
               estimator = "ML")  # ML for maximum likelihood

# Get a summary of the results
summary(cfa_fit, fit.measures = TRUE, standardized = TRUE)
```

## Conducting Comparison CFA

```{r}
# CFA for GAD-7 for comparison
cfa_model2 <-
  'GAD7 =~ Q42_1 + Q42_2 + Q42_3 + Q42_4 + Q42_5 + Q42_6 + Q42_7'

cfa_fit2 <- cfa(cfa_model2, 
               data = CFA_data, 
               estimator = "ML")

# Get a summary of the results
summary(cfa_fit2, fit.measures = TRUE, standardized = TRUE)
```

## Interpretation: Excerpt from Results Section of Thesis

> The final seven-item scale had factor loadings ranging from .60 to .77 and a single factor explaining 47.1% of variance and alpha = .86. CFA was performed on a randomly split half of the sample and indicated excellent fit: CFI = 0.98, TLI = 0.98, RMSEA = 0.07, SRMR = 0.04, χ² (14, 74) = 18.32 (p = 0.192). For comparison, the commonly-used measure of anxiety, the GAD-7 (Spitzer et al., 2006), which is also a seven-item scale, had CFI = 0.98, TLI = 0.97, RMSEA = 0.09, SRMR = 0.05, χ² (14, 72) = 22.40 (p = 0.071), a slightly worse fit, on the same randomly split half of the sample.
