---
title: "Portfolio 10"
---

> The projects should be numbered consecutively (i.e., in the order in which you began them), and should include for each project a description of the goal, the product (computer program, hand graph, computer graph, etc.), the data, and some interpretation. Reports must be reproducible and of high quality in terms of writing, grammar, presentation, etc.
> I am a TA for a Cross-Cultural Psychology course. The students have an assignment in which they must interview someone from a different culture. I've been asked to compile all the demographic information (i.e., where the interviewees are from) and use the data to create interesting and informative visualizations for the class.

## Loading tidyverse
```{r, message=FALSE}
library(tidyverse)
```


## Creating the dataset.
All the data for the dataset is contained on printed out and handwritten demographic forms. I consulted Claude to see how best to create the dataset. After considering the first option (Approach 1 below), I asked about more efficient ways to create the dataset, eventually landing on Approach 3.

### Approach 1.
First, create empty dataset.
```{r}
interviewee_demographics <- tibble(
  age = numeric(),
  gender = factor(levels = c("Male", "Female")),
  race = character(),
  profession_or_major = character(),
  birth_country = character(),
  length_res = numeric(),
  year_moved = numeric()
)
```
Then, edit dataset to add rows for each observation.
```{r}
interviewee_demographics <- interviewee_demographics %>%
  add_row(
    age = 33,
    gender = "Female",
    race = "Black or African American",
    profession_or_major = "Professor",
    birth_country = "Guyana",
    length_res = 2,
    year_moved = 2023
  )
```
Check dataset.
```{r}
interviewee_demographics
```
I then asked Claude is there a way more efficient method of data entry in R, saying that, if not, it would seem best to just create the dataset in Excel and then import it into R.

Claude agreed that using Excel would probably be best, but gave me two other option for quicker data entry in R.

### Approach 2.
```{r}
# Create empty dataframe with your structure
interviewee_demographics <- tibble(
  age = numeric(),
  gender = character(),
  race = character(),
  profession_or_major = character(),
  birth_country = character(),
  length_res = numeric(),
  year_moved = numeric()
)

# Then use this to open a data entry window
interviewee_demographics <- edit(interviewee_demographics)
```

